---
title: "Weekly Homework 3"
author: "Alina Kereszt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/GitHub/Coding/au669161_Kereszt_Alina")
pacman::p_load(tidyverse)
```

## 1) Exercise 1
**Use R to figure out how many elements in the vector below are greater than 2 and then tell me what their sum (of the larger than 2 elements) is.**

```{r}
# create vector
rooms <- c(1, 2, 4, 5, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, 9, 3, NA)

# take length of subsetted vector
# NAs are removed as "greater than 2" is not interpretable on them
length(rooms[!is.na(rooms) & rooms>2])

# take sum of elements in subsetted vector
# NAs are removed as "greater than 2" is not interpretable on them, and R does
# not intake them for numeric operations
sum(rooms[!is.na(rooms) & rooms>2])
```

There are 12 elements greater than 2 and their sum is 55.

## 2) Exercise 2
**What type of data is in the 'rooms' vector?**
```{r}
class(rooms)
```

## Exercise 3
**Submit the following image to Github: Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing**

  **a) the line of code you used to create the object,**

  **b) the 'interviews' object in the Environment, and**

  **c) the file structure of your R project in the bottom right "Files" pane.**

**Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.**

```{r}
interview 
```

























