---
title: "Final project"
author: "Alina Kereszt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, stringr)
```


Data provenance: kaggle.com (https://www.kaggle.com/datasets/jeniagerasimov/ted-talks-info-dataset)

*"TED is a nonprofit devoted to spreading ideas, usually in the form of short, powerful talks (18 minutes or less). TED began in 1984 as a conference where Technology, Entertainment and Design converged, and today covers almost all topics — from science to business to global issues — in more than 100 languages. Meanwhile, independently run TEDx events help share ideas in communities around the world.*

*I enjoy TED talks a lot, they contain important life lessons and often motivate me. So I decided to collect data to analyze it."*

## Dataset preprocessing
I used OpenRefine (CITE) to do a preliminary processing of the data, seeing as it had some inconsistent variables and some variables contained within other variables.

Load data.
```{r}
tedbig <- read.csv("ted.csv")

# for now i'm just taking a lil piece of it bc too slow
ted <- tedbig[1:100,]
```

Clean.
```{r}
# we don't need the hour-minute-second timestamp for when the video was posted,
# so i remove it witha  regex
ted$published_date <- gsub("T[0-9][0-9]:[0-9][0-9]:[0-9][0-9]Z", "", ted$published_date)

# remove unnecessary columns
ted <- ted %>% 
  select(!related_videos) %>% 
  select(!subtitle_languages) %>% 
  select(!youtube_video_code)

# if i wanna look at topics, the topics field is a mess... i'm just gonna pull
# it into its own little dataframe with a little cleaning in case i wanna use
# it
ted$topics <- gsub("[[:punct:]]", "", ted$topics)
ted$topics <- gsub("id[0-9]name", ",", ted$topics)
ted$topics <- gsub("id[0-9][0-9]name", ",", ted$topics)
ted$topics <- gsub("id[0-9][0-9][0-9]name", ",", ted$topics)
ted$topics <- gsub("^.", "", ted$topics)

head(ted)


```























